Study of the effects of magnetic braking on the lithium abundances of the Sun and solar-type stars

### 2020-05-17 project/overleaf/paper.tex
MNRAS revision 2

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{nccmath}
\usepackage{isotope}
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{multicol}
%\usepackage[firstpage]{draftwatermark}
%\SetWatermarkText{DRAFT}

@@

% Abstract of the paper
\begin{abstract}
\textit{Goals.} The study of lithium (Li) surface abundance in the Sun and young stellar globular clusters which are seemingly anomalous in present-day scenarios, as well as the influence of rotation and magnetic braking (MB) on its depletion during pre-main sequence (PMS) and main sequence (MS).
\newline\textit{Methods.} The effects of rotational mixing and of the rotational hydrostatic effects on Li abundances are studied by simulating several grids of PMS \& MS rotating and non-rotating models. The previous effects are combined with the additional impact of the MB (with magnetic field intensities ranging between 3.0 and 5.0 G) and the data obtained from the two sets of simulations are confronted by comparing a number of parameters: Li abundances, angular velocity, effective temperature and size of convective zone.
\newline\textit{Results.} The surface Li abundance for the Sun like models at the end of the PMS and throughout the MS is decreased when rotational effects are included, i.e. the Li depletion rate for rotating models is higher than for non-rotating ones. However, this effect is attenuated when the braking effect produced by a magnetic field is present. This physical phenomena, generally neglected in stellar evolutionary models, impacts also the effective temperature ($\teff$) of the star and therefore its location in the HR diagram. The impact of MB in Li depletion is sensitive to the magnetic field intensity: the higher it is, the lower the Li destruction. A direct link between the magnetic fields and the convective zone (CZ) size is observed: stronger magnetic fields produce shallower CZ's. Another important result of this study is that MB effect must be taken into consideration during PMS if we aim to reproduce Li abundances in young clusters.
\end{abstract}

@@

\section{Introduction} \label{sec_1}
Despite decades of theoretical efforts, a coherent explanation derived from models cannot be found for the Li abundance discrepancies detected in stars belonging to clusters of different ages and which are in the same evolutionary state, either in the PMS or in the MS. Additionally, these theoretical models are not able to explain the abundances detected in the late stages of the MS \citep{Tschape2001}.\par

The denominated standard models, models that include convection only as a mixing process and do not consider other transport options like diffusion or angular momentum loss (AML), have been mainly involved in the elaboration of these predictions \citep{Sestito2005}. Li is destroyed at a temperature $\tli \approx 2.5 x 10^6\; K$ when a Li atom collides with a proton producing two He atoms, something that takes place during proton-proton type II reactions (P-P II), and is therefore directly destroyed in stellar envelopes when the temperature at the base of the convection zone (BCZ) reaches that value. The Sun in particular and the solar-type stars in general are characterized by having a CZ that covers much of the stellar radius during the PMS which causes that their lower limit to exceed $\tli$ \citep{Iben1965}. This depletion stops during the approach to the zero-age MS (ZAMS) when the convection zone retreats and the temperature at BCZ is cooler than $\tli$. In the standard stellar models only the mass and the initial chemical composition determine at what distance from the stellar surface $\tli$ temperature is reached, therefore stars within clusters with similar mass are expected to reach the ZAMS with equal surface Li abundances. Furthermore, they should also show a very similar Li evolution until their approach to the terminal-age MS (TAMS). During this same period the convection mechanisms unleash a mixing process that homogenizes the chemical composition of the convective envelope, from its lower limit till the star surface. However, different abundances of Li have been observed for different star populations \citep[][and references therein]{Somers2014}.\par

Theoretical models of stars do not inform about the initial amount of Li that a star has, they only describe how it is exhausted. Therefore, in order to make an accurate estimate of the initial abundance of Li, it is a prerequisite to be able to compare observations and models beforehand. Our Sun represents a unique exception, since it allows us to know the current abundance of this element in its photosphere, $A(\isotope[7]{Li}) = 1.1 \pm 0.1 \, dex$ \citep{Jeffries2004}, where $A(\isotope[7]{Li})$ is defined according to Eq.~\ref{eq:A_Li}\par

@@

On the other hand, the initial abundance of $A(\isotope[7]{Li}) = 3.34 \, dex$ is obtained from meteorite measurements \citep{Randich2006}. According to the theory for newly born stars, we have that the initial abundance of Li can be estimated fairly accurately from photospheric measurements on T-Tauri type stars, or from the hottest F stars forming part of slightly older clusters. For the latter, the current theory of stellar evolution suggests that Li must not yet be exhausted. The results obtained from the measurements on both types of stars allow us to fix the initial abundance of Li in the interval $3.0 \, dex < A(\isotope[7]{Li}) < 3.4 \, dex$ \citep{Randich2006}.\par

The impact of rotation both on PMS and Li depletion for solar-type stars has already extensively debated in the past \citep{Pinsonneault1997,Jeffries2004,Somers2014} and revised more recently on the basis of the availability of more accurate measures \citep{Bouvier2016}. However, these previous studies focused mainly on the hydrostatic effects and did not consider the influence of a coupling between the star magnetic field and its possible spin-down effect. It is thought that AML has a direct influence in the mixing processes and that transport of the momentum could be produced by a number of relevant mechanisms: mass loss, magnetic fields and gravity waves (g-modes). Concerning the two latter, gravity waves \citep{Charbonnel2005} and magnetic fields \citep{Eggenberger2009} have the property of transmitting angular momentum (AM) much more effectively than inducing mixing \citep{Denissenkov2007}. As a consequence of this increment in the AM transport efficiency, the amount of differential rotation between the radiative and the convective zones of the star is reduced (a solid-body rotation is fostered), as are the induced rotational instabilities. Additionally, magnetic fields would also originate in the limits between these zones, in the so-called tachoclines \citep{Aschwanden2014, Guerrero2016}, which interact with the solar wind particles that are forced to rotate in sync with it and therefore slow down the star, i.e. the magnetic braking (MB) effect. \par

Today, standard models are not able to replicate in a satisfactory way the observed values of Li abundance on star surfaces with the predictions that these models yield. This lack of agreement makes us think that certain physical mechanisms that influence the destruction of Li are either being modelled improperly or simply not being taken into account, e.g. the magnetic braking. In view of these findings, it seems evident that there are still open questions to be resolved on the subject of the processes that participate, directly or indirectly, in the mixing mechanisms that are not incorporated in the standard stellar models. Therefore, it is essential a proper consideration of the interactions between rotation and magnetic fields when it comes to study the AM distribution. \par

\subsection{Effects of magnetic fields}
The origin of magnetic fields in the stars is still unknown. The organized magnetic fields detected in some O stars \citep{Wade2010} might be fossil fields \citep[see][for details ]{Dudorov2014}, or fields produced through a dynamo mechanism \citep{Cantiello2009}. Besides that, the presence of magnetic fields with intensities of the order of kG \citep{Hussain2014} has been a sine qua non condition for dealing with certain data observed in T Tauri type stars in young, accreting PMS systems \citep{Johns-Krull2007}.\par

The mass of stars plays a determining role in the existence or not of an external convective envelope. In turn, the presence of this convective region conditions the evolution of magnetic fields. Massive stars do not have external convective zones although they can expose a convective core. On the other hand the less massive stars, such as the Sun, have convective covers and a radiative core. In both types of stars strong magnetic fields can be generated in the tachocline but with notable differences \citep[e.g.][for details]{Chabrier2006,Charbonneau2010}.\par

\subsubsection{Internal Magnetic Fields}
Our Sun is an active star as far as the magnetic field is concerned. According to the theoretical models this magnetic field is generated by a dynamo in the tachocline and with an intensity about $B\approx10^{5}\,\Gauss$ \citep{Maeder2003a,Dudorov2014}. At the moment of their generation, the lines of that magnetic field will have a poloidal orientation and will be affected by the differential rotation and forced to adopt a toroidal typology (the $\omega$-effect). On the other hand, the opposite case can also occur, starting from a toroidal magnetic field to obtain a poloidal configuration. This would be possible thanks to the convective movements of material that are produced in the CZ of the star and to assume that this material behaves as a perfect conductor. This second condition allows to assume that the magnetic field lines are "frozen" in the conductive fluid and therefore must move in solidarity with it ($\alpha$-effect).\par

\subsubsection{Surface Magnetic Fields} \label{surf_mf}
The intensity of the magnetic field measured on the Sun's surface is about $B\approx1\,\Gauss$ \citep{Weber1967,DAntona2000,Morin2012}, twice the value of the Earth's one. Additionally its topology is varied and complex. The dynamo model establishes that the magnetic field produced inside the star reaches the surface with the help of convective processes taken place in the CZ and gives rise to magnetically active regions on the solar surface. Besides, rotating stars that expose a significant outer CZ can produce surface magnetic fields through a dynamo mechanism \citep[e.g.][]{Brandenburg2004,Charbonneau2010,Brun2017}. Whatever the origins of surface magnetic fields are, it is expected to couple to the wind mass-loss and, if this is strong enough, to produce MB \citep[e.g.][]{UdDoula2002,Ud-Doula2007,Ud-Doula2008,Meynet2010}.\par

\subsection{Effects of magnetic braking}
In its most widely accepted approach, the MB is linked to Skumanich's work \citep{Skumanich1972} in which he develops an empirical law of its evolution in parallel to the age of the star. To calibrate this law, Skumanich relied on data obtained from type G stars found in the MS. The influence of the MB on the evolution of the star is closely related to the transport of the AM. According to \citet{Meynet2010}, two main mechanisms are distinguished attending to how the transport of the AM is produced:

\begin{enumerate}
    \item Differential rotation: the AM transport is pushed by southern currents and shear instabilities.
    \item Solid body rotation: when the transport of AM is very efficient, solid body rotation is 
    maintained throughout the entire MS phase.
\end{enumerate}

The AML caused by MB depends directly on the amount of mass lost by the star due to the stellar winds. The estimated mass loss ratio for a solar-type star, approximately $10^{-14}\msun \; yr^{-1}$ \citep{Noerdlinger2008}, and the resulting AML derived from that value can be considered relatively modest as to decisively influence the evolution of the star. Furthermore, the combined effect of convective movements and differential rotation lead to the generation of magnetic fields. Those magnetic fields end up reaching the surface of the star, so that the AML can be increased in several orders of magnitude \citep{Langer2012}. That increment is a consequence of the magnetic field that forces the stellar wind ionized particles to turn with the same angular speed, winds which extends a distance several times the radius of the star \citep[see][for more details]{UdDoula2002,Ud-Doula2007,Ud-Doula2008}.

\section{Method} \label{sec_2}

In the remainder of this paper, we describe a physically sound yet computationally simple model which will be used as an extension to Modules for Experiments in Stellar Astrophysics stellar evolution code (MESA; \citeauthor{Paxton2011} \citeyear{Paxton2011}, \citeyear{Paxton2013}, \citeyear{Paxton2015}, \citeyear{Paxton2018}, \citeyear{Paxton2019}) for the calculation of the AML as a result of the torque applied by a magnetically-coupled stellar winds.

\subsection{Modelling magnetic braking} \label{mod_mb}
We'll start enumerating the most relevant aspects and assumptions made in modelling the evolution of rotation, magnetic braking and angular momentum in MESA for the calculation of the AML as a result of the torque applied by a magnetically-coupled stellar winds.\par

In the presence of mass-loss, AM is removed by the material in the stellar wind, resulting in a spin-down effect. This is quite important for example in massive stars, that are known to be rapidly rotating and have strong stellar winds. It has been also observed that strongly magnetic intermediate-mass stars typically have rotation rates much slower than other stars in their parent population \citep{Mathys2006}. In those stars, the presence of such magnetic field will interact with the mass loss and the Alfv\'{e}n radius ($R_{A}$) plays an important role in this interaction.\par

@@

Making use of $\eta_*$, the $R_{A}$ is defined as the point in which the magnetic field energy density $B^{2}/8\upi$, marked by magnetic field strength ($B$), and the kinetic energy density $\frac{1}{2}\rho\nu^{2}$, defined by the mass density ($\rho$) and velocity of the stellar wind $(\nu)$, are balanced. That is, where the coefficient between both is equal to 1. In case that $R_{A}$ is greater than the star radius, then the wind flow will have to follow the magnetic field. As a consequence, the material will leave the stellar surface with a higher specific AM, as the co-rotation radius has increased and it will roughly correspond to the $R_{A}$. This effect is known as MB.\par

Each of these factors has a decisive influence on the future of the star and therefore in our routine implementation. On the one hand, stars with similar initial masses but different mass loss ($\Dot{M}$) ratios will end up evolving very differently. On the other hand, the ionized particles carried by the solar wind not only contribute to the mass loss but also to the loss of kinetic ($K_e$) energy that is deposited in the interstellar medium. Given a star with a spherically symmetric wind, $\Dot{M}$ and $K_e$ are characterized and related by the following expressions:

\begin{ceqn}
\begin{align}
    \Dot{M} &= 4\upi r^2\rho\nu \label{eq:mass_loss}\\
    K_e &= 0.5\Dot{M}\nu_{esc}
\end{align}
\end{ceqn}

@@

In turn, of the observable characteristics of the solar wind, we especially care about the parameters related to the amount of mass that is able to pull out of the outer layers of the star in a given time interval, and the speed that reaches the wind itself measured at a great distance from the star, the terminal velocity ($\nu_\infty$). In general, a magnetically channeled outflow will have a complex stream geometry but for convenience, the Eq.~\ref{eq:mass_loss} simply characterizes the wind strength in terms of a spherically symmetric mass-loss rate. $\nu$ can be characterized by the radial variation of outflow velocity in terms of the velocity law (Eq.~\ref{eq:vel_law}), where $r$ represents the distance from the center of the star to the point where the velocity is to be measured, and $R_*$ is the radius of the star:
\begin{ceqn}
\begin{equation}
    \nu(r) = \nu_\infty (1-R_*/r) \label{eq:vel_law}
\end{equation}
\end{ceqn}
where $\nu_\infty$ is the terminal wind velocity defined as the velocity that the wind, the outflowing matter, reaches at large distance from the central star, where it is not accelerated anymore by the wind driving force but its deceleration due to interaction with the interstellar medium (ISM) is negligible \citep{Niedzielski2002}.\par

The line-driven winds of massive OB stars have terminal velocities that scale with the photospheric escape velocity \citep{Lamers2000} according to:
\begin{ceqn}
\begin{align}
\nu_{esc} &= \sqrt{\frac{2GM_*}{R_*}} \label{eq:vesc} \\
\nu_\infty &\simeq 1.92 \;\nu_{esc}\\
\nu_\infty &\simeq 1.92 \; x \; 618 \; \Bigg(\sqrt{\frac{R_{\sun}}{R_*}\frac{M_*}{\msun}} \;\Bigg) \label{eq:vinf}
\end{align}
\end{ceqn}
where Eq.~\ref{eq:vesc} describes the Newtonian escape velocity from the stellar surface, $G$ is the gravitational constant and Eq.~\ref{eq:vinf} represents the terminal velocity in terms of escape velocity, radius and mass of the Sun.\par

Given that the version of MESA used in this work does include neither the treatment of surface magnetic fields (it does account for transport by internal magnetic fields of AM and chemical elements) and therefore, nor for magnetic braking effects, it is necessary to extend the simulator to include this phenomenon. In the present work we have followed the theoretical work on MB in the Sun carried out by \citet{Weber1967}, making use of Eq.~ \ref{eq:j_dot} for the calculation of the AML.
\begin{ceqn}
\begin{equation}
 \Dot{J} = \frac{2}{3} \Dot{M}\Omega R^{2}_{A} \label{eq:j_dot}
\end{equation}
\end{ceqn}
where $\Dot{M}$ is the mass loss rate, $\Omega$ the angular velocity at star surface and $R_A$ the Alfv\'{e}n radius. \par

The above expression can be rewritten so that it depends on the $\eta_*$ instead of the $R_A$ as follows:
\begin{ceqn}
\begin{equation}
 \Dot{J} = \frac{2}{3} \Dot{M}\Omega R^{2}_{*}\eta_* \label{eq:j_dot_mesa}
\end{equation}
\end{ceqn}
which is a more convenient expression to be implemented in MESA because is based on values directly exposed during the simulations and suitable for the approach described in \citet{Ud-Doula2007}.

@@

\subsection{Mesh Models} \label{sec_mesh}

In its simplest form the implementation of the model requires the specification of both the surface magnetic field strength ($B$) and the initial rotation rate ($\Omega$) at star surface. The numerical simulations trace the rotational history and A(Li) of a $1\, \msun$ star for a variety of initial values for $B$ and $\Omega$. Although stars have a three-dimensional structure and the equations that model their evolution must take this fact into account, it has been proven sufficient to solve the equations in only one dimension, extrapolating the results to the other two \citep[shellular approximation,][]{Meynet1997}. MESA uses this simplification to accelerate computational calculations \citep{Paxton2013}.\par

\begin{table}
	\centering
	\caption{Summary of adopted physics in MESA \citep[based on][]{Choi2016}.}
	\label{tab:phy_mesa}
	\begin{tabular}{ll} 
		\hline
		Parameter & Adopted prescriptions and values\\
		\hline
		Solar Abundance & $X_{\odot}=0.7154, Y_{\odot}=0.2703, Z_{\odot}=0.0142$\\
		Equation of State & OPAL+SCVH+MacDonald+HELM+PC\\
		Opacity & OPAL Type I for log T $\geq$ 4 \\ & Ferguson for logT $\leq$ 4\\
		Reaction Rates & JINA REACLIB\\
		Boundary Conditions & ATLAS12; $\tau$=100 tables + photosphere\\
		Difussion & Track \isotope[1]{H}, \isotope[2]{He}, \isotope[7]{Li}, \isotope[7]{Be}\\
		Rotation & Differential rotation at PMS \& MS\\
		Convection & MLT + Ledoux, $\alpha_{MLT}$ = 1.82\\
		Overshoot & time-dependent, diffusive, \\ & $f_{ov,core}=0.0160$,\\ 
		& $f_{ov,sh}=0.0174$\\
		Semiconvection & $\alpha_{sc}=0.1$\\
		Thermohaline & $\alpha_{th}=666$\\
		Rotational Mixing & Include SH, ES, GSF, SSI \& DSI\\
		Magnetic Effects & Magnetic braking based on idealized \\ & monopole field\\
		Magnetic Field & B(G) variable between [3.0 - 5.0]\\
		Mass Loss & activated, $\Dot{M}_{max} = 10^{-3} \: \msun \: yr^{-1}$\\
		Angular Moment Loss & activated, $\Dot{J} = \frac{2}{3} \Dot{M}\Omega R^{2}_{A}$\\
		\hline
	\end{tabular}
\end{table}

We define here also additional conventions and assumptions adopted in the simulations. For the models presented in this paper we adopt solar-scaled abundances and assumes the \citet{Asplund2009} protosolar birth cloud metallicity ($Z = Z_{\odot, \mathrm{pr} = 0.0142}$) not the current photospheric metallicity ($Z = Z_{\odot, \mathrm{ph} = 0.0134}$) as the reference value. We also adopt the following nominal values to express stellar properties values in SI units $\rsun = 6.957x10^{10}\, cm$ and $\msun = 1.988x10^{33}\, g$ which are consistent with IAU resolutions \citep{Mamajek2015}. For a detailed description of the physics adopted in this paper, refer to \citet{Choi2016}. That work, in particular its section 4.1, has been used as a starting point for ours regarding the parameterization of the MESA project which is calibrated to reproduce the measured element abundances on the solar surface.\par

The models include already rotation during the PMS as there evidences that advocate for a strong established relationship between Li destruction and rotation on that phase (\citeauthor{Bouvier2016} \citeyear{Bouvier2016}, \citeyear{Bouvier2018}), and by activating it in the MESA models, the evolutionary computational code takes into account the following effects:

\begin{enumerate}
    \item AM transport from the radiative interior to the convective envelope in response to the rotational deceleration of the star surface layers\label{itm:1}.
    \item AM redistribution associated with changes in internal structure during the process of contraction to the MS\label{itm:2}.
    \item AML as a result of the torque applied to the convection zone by a magnetically coupled wind\label{itm:3}.
\end{enumerate}

The enumerated effects \ref{itm:1} and \ref{itm:2} are standard features offered by MESA. On the contrary, the effect \ref{itm:3} is not available as an out-of-the-box feature during the elaboration of this paper and represents our extension to MESA. Additionally, we do not take into account either the influence of internal magnetic fields or their existence during the T-Tauri phase. We adopt a simple and pragmatic theoretical approach to establish their occurrence when the star is reaching the ZAMS phase. The adopted criterion is based on the simultaneous existence of an extensive convective layer and a radiative core. From this moment on, the MB routine will be activated, acting as an additional mechanism to those existing in the MESA evolutionary code that participates in the star AML. Additionally, we assume that the magnetic field does not vary its intensity throughout the evolution of the star until it reaches the TAMS.\par

>>>>

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{nccmath}
\usepackage{isotope}
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{multicol}
\usepackage[normalem]{ulem}
%\usepackage[firstpage]{draftwatermark}
%\SetWatermarkText{DRAFT}

@@

% Abstract of the paper
\begin{abstract}
\textit{Goals.} The study of lithium (Li) surface abundance in the Sun and young stellar globular clusters which are seemingly anomalous in present-day scenarios, as well as the influence of rotation and magnetic braking (MB) on its depletion during pre-main sequence (PMS) and main sequence (MS).
\newline\textit{Methods.} The effects of rotational mixing and of the rotational hydrostatic effects on Li abundances are studied by simulating several grids of PMS and MS rotating and non-rotating models. The previous effects are combined with the additional impact of the MB, (with magnetic field intensities ranging between 3.0 and 5.0 G). The data obtained from the two sets of simulations are confronted by comparing a number of parameters: Li abundances, angular velocity, effective temperature and size of convective zone.
\newline\textit{Results.} The surface Li abundance for the Sun like models at the end of the PMS and throughout the MS is decreased when rotational effects are included, i.e. the Li depletion rate for rotating models is higher than for non-rotating ones. However, this effect is attenuated when the braking effect produced by a magnetic field is present. This physical phenomena, generally neglected in stellar evolutionary models, impacts also the effective temperature ($\teff$) of the star and therefore its location in the HR diagram. The impact of MB in Li depletion is sensitive to the angular speed and this in turn on the magnetic field intensity: the higher the latter, the lower the Li destruction. A direct link between the magnetic fields and the convective zone (CZ) size is observed: stronger magnetic fields produce shallower CZ's. This result suggests that MB effect must be taken into consideration during PMS if we aim to reproduce Li abundances in young clusters.
\end{abstract}

@@

\section{Introduction} \label{sec_intro}
Despite decades of theoretical efforts, a coherent explanation for the Li abundance discrepancies detected in star clusters cannot be found. In young and old clusters these differences have been documented for stars in both PMS and MS  evolutionary stages. Additionally, these theoretical models are not able to explain the abundances detected in the late stages of the MS \citep{Tschape2001}.\par

Standard models, models that include convection only as a mixing process and do not consider other transport options like diffusion or angular momentum loss (AML), have been mainly involved in the elaboration of these predictions \citep{Sestito2005}. Li is destroyed at a temperature $\tli \approx 2.5 x 10^6\; K$ when a Li atom collides with a proton producing two He atoms. That takes place during proton-proton type II reactions (P-P II), and is therefore directly destroyed in stellar envelopes when the temperature at the base of the convection zone (BCZ) reaches  $\tli$. The Sun in particular and the solar-type stars in general are characterized by having a CZ that covers much of the stellar radius during the PMS which causes that their lower limit to exceed $\tli$ \citep{Iben1965}. This depletion stops during the approach to the zero-age MS (ZAMS) when the convection zone retreats and the temperature at BCZ is cooler than $\tli$. In the standard models only the mass and the initial chemical composition determine at what distance from the stellar surface $\tli$ temperature is reached. Therefore stars with similar mass belonging clusters are expected to reach the ZAMS with equal surface Li abundances. Furthermore, they should also show a very similar Li evolution until their approach to the terminal-age MS (TAMS). During this period the convection mechanisms unleash a mixing process that homogenizes the chemical composition of the envelope, from its lower limit till the star surface. However, different abundances of Li have been observed for different stellar populations \citep[see][and references therein]{Somers2014}.\par

Theoretical models take initial Li abundance as an input parameter and only describe how it is exhausted. Therefore, in order to make an accurate estimate of the initial abundance of Li, it is a prerequisite to be able to compare observations and models beforehand. Our Sun represents a unique exception, since it allows us to know the current abundance of this element in its photosphere, $A(\isotope[7]{Li}) = 1.1 \pm 0.1 \, dex$ \citep{Jeffries2004}, where $A(\isotope[7]{Li})$ is defined

@@

and its initial abundance of $A(\isotope[7]{Li}) = 3.34 \, dex$ is obtained from meteorite measurements \citep{Randich2006}. For other stars the initial Li abundance is found to be in the interval $3.0 \, dex < A(\isotope[7]{Li}) < 3.4 \, dex$. This has been estimated from  \citet{Randich2006}.\par

The impact of rotation both on PMS and Li depletion for solar-type stars has already extensively debated in the past \citep{Pinsonneault1997,Jeffries2004,Somers2014} and revised more recently on the basis of the availability of more accurate measures \citep{Bouvier2016}. However, these previous studies focused mainly on the hydrostatic effects and did not consider the influence of a coupling between the stellar magnetic field and its possible spin-down effect. AML has a direct influence in the mixing processes. It can be produced by a number of relevant mechanisms: mass loss, magnetic fields and gravity waves (g-modes). Gravity waves \citep{Charbonnel2005} and magnetic fields \citep{Eggenberger2009} transmit angular momentum (AM) much more effectively than inducing mixing \citep{Denissenkov2007}. As a consequence, the amount of differential rotation between the radiative and the convective zones of the star is reduced. Additionally, magnetic fields would also originate in the limits between these zones, in the so-called tachoclines \citep{Aschwanden2014, Guerrero2016}, which interact with the solar wind particles that are forced to co-rotate with it. This slows down the star, effect which is known as magnetic braking (MB). \par

Today, standard models are not able to replicate the observed values of Li abundance on stellar surfaces. This might happen because certain physical mechanisms that influence the destruction of Li are either being modelled improperly or simply not being taken into account, e.g. the magnetic braking. In this paper we study the interactions between rotation and magnetic fields when it comes to discuss the AM distribution. \par

\section{Method} \label{sec_method}

In the remainder of this paper, we describe a semi-empirical approach for the calculation of the AML as a result of the torque applied by a magnetically-coupled stellar winds. This has been implemented as an extension to Modules for Experiments in Stellar Astrophysics stellar evolution code \citep[MESA; ][]{Paxton2011, Paxton2013,Paxton2015, Paxton2018, Paxton2019}.

\subsection{Modelling magnetic braking} \label{mod_mb}
We start by enumerating the most relevant aspects and assumptions made in modelling the evolution of rotation, magnetic braking and angular momentum in MESA. These will be considered for the calculation of the AML as a result of the torque applied by a magnetically-coupled stellar winds.\par

@@

where $B^{2}/8\upi$ is the magnetic field energy density, $\frac{1}{2}\rho\nu^{2}$ the kinetic energy density, $B$ the magnetic field intensity, $\rho$ the mass density, and $\nu$ the stellar wind velocity.\par

Stars with similar initial masses but different mass loss ($\Dot{M}$) ratios will end up evolving very differently. The ionized particles carried by the solar wind not only contribute to the mass loss but also to the loss of kinetic ($K_e$) energy that is deposited in the interstellar medium. Given a star with a spherically symmetric wind, $\Dot{M}$ is characterized by the following expression:

\begin{ceqn}
\begin{equation}
    \Dot{M} = 4\upi r^2\rho\nu \label{eq:mass_loss}
\end{equation}
\end{ceqn}

@@

In general, a magnetically channeled outflow has a complex stream geometry but for convenience, the Eq.~\ref{eq:mass_loss} simply characterizes the wind strength in terms of a spherically symmetric mass-loss rate. $\nu$ can be characterized by the radial variation of outflow velocity in terms of the velocity law (Eq.~\ref{eq:vel_law})
\begin{ceqn}
\begin{equation}
    \nu(r) = \nu_\infty (1-\rstar/r) \label{eq:vel_law}
\end{equation}
\end{ceqn}
where $r$ represents the distance from the center of the star to the point where the velocity is to be measured, $R_*$ is the radius of the star, and $\nu_\infty$ is the terminal wind velocity. $\nu_\infty$ is defined as the velocity that the wind reaches at large distance from the central star, where it is not accelerated anymore by the wind driving force but its deceleration due to interaction with the interstellar medium (ISM) is negligible \citep{Niedzielski2002}.\par

The line-driven winds of OB stars have terminal velocities that scale with the photospheric escape velocity \citep{Lamers2000} according to:
\begin{ceqn}
\begin{align}
    \nu_\infty &\simeq 1.92 \;\nu_{{\rm esc}} \label{eq:vinf}\\
    \nu_{{\rm esc}} &= \sqrt{\frac{2\,G\,\mstar}{\rstar}} \label{eq:vesc}
\end{align}
\end{ceqn}
where Eq.~\ref{eq:vesc} describes the Newtonian escape velocity from the stellar surface, where $G$ is the gravitational constant,   ${\rstar}$ the radius, and ${\mstar}$ the mass of the star.

Making use of (\ref{eq:vesc}), $\nu_\infty$ can be reformulated as:

\begin{ceqn}
\begin{equation}
    \nu_\infty \simeq 1.92 \; x \; 618 \; \Bigg(\sqrt{\frac{\rsun}{\rstar}\frac{\mstar}{\msun}} \;\Bigg) \label{eq:vinf}
\end{equation}
\end{ceqn}
where ${\rsun}$ is the radius, and ${\msun}$ the mass of the Sun.

It has been observed that strongly magnetic intermediate-mass stars typically have rotation rates much slower than other stars in their parent population \citep{Mathys2006}. In those stars, the magnetic fields interacts with the mass loss, where the Alfv\'{e}n radius ($R_{A}$) plays an important role. Making use of $\eta_*$ (Eq. \ref{eq:wind_conf}), $R_{A}$ is defined as the point in which the magnetic field energy density and the kinetic energy density are balanced. In case that $R_{A}$ is greater than the stellar radius, then the wind flow will have to follow the magnetic field. As a consequence, the material leaves the stellar surface with a higher specific AM, as the co-rotation radius has increased and it roughly corresponds to $R_{A}$. Following \citet{Weber1967} the AML can be calculated by:
\begin{ceqn}
\begin{equation}
 \Dot{J} = \frac{2}{3} \Dot{M}\Omega R^{2}_{A} \label{eq:j_dot}
\end{equation}
\end{ceqn}
where $\Dot{M}$ is the mass loss rate, $\Omega$ the angular velocity at star surface and $R_A$ the Alfv\'{e}n radius. \par

The above expression can be rewritten according to \citet{Ud-Doula2008} so that it depends on the $\eta_*$ instead of the $R_A$ as follows:
\begin{ceqn}
\begin{equation}
 \Dot{J} = \frac{2}{3} \Dot{M}\Omega R^{2}_{*}\eta_* \label{eq:j_dot_mesa}
\end{equation}
\end{ceqn}
which is a more convenient expression to be implemented in MESA because is based on values directly exposed during the simulations.

@@

\subsection{Mesh Models} \label{sec_mesh}
According to the formulation in the previous sections, the only two free parameters of our implementation are $B$ and $\Omega$. The numerical simulations traced the rotational history and $A(\isotope[7]{Li})$ of a $1\, \msun$ star for a variety of initial values for $B$ and $\Omega$ (see also Table \ref{tab:phy_mesa}). MESA uses the shellular approach \citep{Meynet1997} to account for hydrostatic effects of rotation into 1D stellar models.\par

\begin{table}
	\centering
	\caption{Summary of adopted physics in MESA \citep[based on][]{Choi2016}.}
	\label{tab:phy_mesa}
	\begin{tabular}{ll} 
		\hline
		Parameter & Adopted prescriptions and values\\
		\hline
		Solar Abundance & $X_{\odot}=0.7154, Y_{\odot}=0.2703, Z_{\odot}=0.0142$\\
		Equation of State & OPAL+SCVH+MacDonald+HELM+PC\\
		Opacity & OPAL Type I for log T $\geq$ 4 \\ & Ferguson for logT $\leq$ 4\\
		Reaction Rates & JINA REACLIB\\
		Boundary Conditions & ATLAS12; $\tau$=100 tables + photosphere\\
		Difussion & Track \isotope[1]{H}, \isotope[2]{He}, \isotope[7]{Li}, \isotope[7]{Be}\\
		Rotation & Differential rotation at PMS \& MS\\
		Convection & MLT + Ledoux, $\alpha_{{\rm MLT}}$ = 1.82\\
		Overshoot & time-dependent, diffusive, \\ & $f_{{\rm ov,core}}=0.0160$,\\ 
		& $f_{{\rm ov,sh}}=0.0174$\\
		Semiconvection & $\alpha_{{\rm sc}}=0.1$\\
		Thermohaline & $\alpha_{{\rm th}}=666$\\
		Rotational Mixing & Include SH, ES, GSF, SSI \& DSI\\
		Magnetic Effects & Magnetic braking based on idealized \\ & monopole field\\
		Magnetic Field & B(G) variable between [3.0 - 5.0]\\
		Mass Loss & activated, $\Dot{M}_{{\rm max}} = 10^{-3} \: \msun \: yr^{-1}$\\
		Angular Moment Loss & activated, $\Dot{J} = \frac{2}{3} \Dot{M}\Omega R^{2}_{A}$\\
		\hline
	\end{tabular}
\end{table}

We adopted solar-scaled abundances and assume \citet{Asplund2009} solar initial abundance ($Z = Z_{\odot, \mathrm{pr} = 0.0142}$). We also adopted the following nominal values to express stellar properties values in SI units, $\rsun = 6.957x10^{10}\, cm$ and $\msun = 1.988x10^{33}\, g$ which are consistent with IAU resolutions \citep{Mamajek2015}. For a detailed description of the physics adopted in this paper, we refer the reader to \citet{Choi2016}. That work was used as starting point for ours regarding the parameterization of the MESA project which was calibrated to reproduce the measured element abundances on the solar surface.\par

The models included rotation during the PMS as there are evidences that advocate for a strong established relationship between Li destruction and rotation on that phase (\citeauthor{Bouvier2016} \citeyear{Bouvier2016}, \citeyear{Bouvier2018}). We modeled rotation such as AML is computed as a result of the torque applied to the convection zone by a magnetically-coupled wind. Additionally, we did not take into account either the influence of internal magnetic fields or their existence during the T-Tauri phase. We adopted a simple and pragmatic theoretical approach to establish when the star is reaching the ZAMS phase. The selected criterion was based on the simultaneous existence of an extensive convective layer and a radiative core. From this moment on, the MB routine was activated, acting as an additional mechanism to those existing in the MESA evolutionary code that participated in the star AML. We assumed that the magnetic field did not vary its intensity throughout the evolution of the star until it reached the TAMS. In addition, other rotation effects were also considered during the evolution of the models: the AM transport from the radiative interior to the convective envelope and the AM redistribution associated with changes in the internal structure during the process of contraction to the MS.\par

In this study we focused on the indirect role played by the MB on the Li destruction. This is a consequence of its influence on the rotational history of solar-type stars. Our major objective was to check how MB could contribute to explain the evolution of Li in the Sun and other solar-type stars. A set of different scenarios with magnetic field strengths ranging between 3.0 and 5.0 G were exercised. We computed the evolution of $1\,\msun$ stellar models at solar initial metallicity with $\oomegac$ varying between $0.0084$ and $0.0336$. More advanced models for more complex arrangements of the magnetic field as well as the evolution of its intensity throughout the life of the star were left for later work.\par
